RNA-seq from 4 cell populations

### 2021-08-04 deconvolution/fans4/molec_profiling.Rmd
bug fix

infoExpand = rbind(data.frame(info[,c(1,3)], dataset="FANS4", stringsAsFactors=FALSE), 
  data.frame(Name=colnames(gtex_count[,id]), CellType = "GTEx BA9", dataset="GTEx", stringsAsFactors=FALSE))
rownames(infoExpand) = infoExpand$Name

# the first PC separates FANS from bulk, so subtract that and use residuals
fit = dream(geneExprCombined, ~ dataset, infoExpand)
resid = residuals(fit, geneExprCombined)

@@

xlab = paste0('PC1 (', format(varFrac[1]*100, digits=3), '%)')
ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ggplot(df_pca, aes(PC1, PC2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab(xlab) + ylab(ylab)

# xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
# ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
# ggplot(df_pca, aes(PC2, PC3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab(xlab) + ylab(ylab)

@@

```{r GTEX.MDS}

d = as.dist(1 - cor(geneExprCombined)^2)
# d = as.dist(1 - cor(apply(geneExprCombined, 2, RNOmni::rankNorm), method="sp")^2)

df_mds = data.frame(cmdscale( d, k=3))

@@

ggplot(df_mds, aes(X1, X2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab('Coordinate 1') + ylab('Coordinate 2')

ggplot(df_mds, aes(X2, X3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab('Coordinate 2') + ylab('Coordinate 3')
```

@@

```{r exit, cache=FALSE}
knitr::knit_exit()
```

>>>>

infoExpand = rbind(data.frame(info[,c("Name", "CellType", "Batch")], 
  dataset="FANS4", stringsAsFactors=FALSE), 
  data.frame(Name=colnames(gtex_count[,id]), CellType = "GTEx BA9", dataset="GTEx", Batch = "GTEx", stringsAsFactors=FALSE))
rownames(infoExpand) = infoExpand$Name

# the first PC separates FANS from bulk, so subtract that and use residuals
fit = dream(geneExprCombined, ~ Batch, infoExpand)
resid = residuals(fit, geneExprCombined)

@@

xlab = paste0('PC1 (', format(varFrac[1]*100, digits=3), '%)')
ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ggplot(df_pca, aes(PC1, PC2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, 'GTEx BA9' = "grey")) + xlab(xlab) + ylab(ylab)

xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
ggplot(df_pca, aes(PC2, PC3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, 'GTEx BA9' = "grey")) + xlab(xlab) + ylab(ylab)
```

@@

```{r GTEX.MDS}
d = as.dist(1 - cor(resid, method="spearman")^2)
# d = as.dist(1 - cor(apply(geneExprCombined, 2, RNOmni::rankNorm), method="sp")^2)

df_mds = data.frame(cmdscale( d, k=2))

@@

ggplot(df_mds, aes(X1, X2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, 'GTEx BA9' = "grey")) + xlab('Coordinate 1') + ylab('Coordinate 2')

# ggplot(df_mds, aes(X2, X3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, , 'GTEx BA9' = "grey")) + xlab('Coordinate 2') + ylab('Coordinate 3')
```

@@
